% Homework #2 
% COSC552 HCI
%
% Byron Heads
% E00062946

\documentclass[12pt]{article}

\title{Homework Set 2 \\
    COSC552 HCI}
\author{ Byron Heads \\
    E00062946 }
\date{\today}

\begin{document}
\maketitle

\section*{A}

\subsection*{Development House}
A commercial development house can have the resources to hire personnel
that are knowledgeable in the field.  This gives
them the ability to add knowledge in different areas.  Many development 
houses also employ requirements engineers to
gain knowledge of their potential user base.  This level of detail can
also lead to problems; software produced maybe too specialized and
become difficult to maintain or adapt to changes in customer needs.
When dealing with tool selection a development house often chooses
to work with a specific technology set.  This limitation may lead to
using a language, library, or technology that may not
be a good fit for the problem domain.  Though this may have limitations,
it can still be a positive situation for the development of specific  
software.  The developers will know what their tool selection can do
and how to do it.

When designing a user interface the development house often has in house
design rules and guidelines that its developers follow.  The large number
of experienced developers will also be able to apply their own experience 
to the user interface design and to the software prototype.  Though the 
development house may be able to design a consistent and usable interface,
they rarely are able to develop outside of these constraints.  This can
be due to the libraries and tools used, or company rules and policies. 

\subsection*{College Students}

A group of college students will have less development experience then
then the development house.  This will lead to less effective design and
code, but the college students are not limited to tools and technology 
that maybe used at a development house.  

College students will have less experience in user interface design and 
will draw design rules from their own experience in using other interfaces.
College students will be focusing on the knowledge and experience they
will gain by developing a large commercial project.  Students often use 
trial and error when writing software.  This will lead to more
experimentation in the interface design.   

\section*{B}

There are many tool to use in designing a system to support users with 
different disabilities.  Before adding these tools you need to have user
controls that enable or disable different tools that depend on the user
that is logged in to the system.  This prevents the user from having to
use the program's options menu to activate or deactivate the different
tools.

To assist a deaf user the interface needs to ensure that any audible signal
has an equivalent visual signal on screen.  This can include flashing 
icons, shaking windows, or in important situations, a pop-up window can
be used.  These visual signals can be used with the audible signals for
all users.

Users with manual dexterity problems can be helped with a tool that
increases the size of on-screen controls and fonts.  The can make clicking
buttons, selecting menus, or selecting text easier.  A system can also
be deployed with a touch screen which may make it easier for some users.

The interface can also include tools to assist users with vision 
impairment by giving the user the option to increase the size of the font
and controls.  The interface can use specific operating system set color
scheme, such as high contrast colors to help users.  The interface can 
take advantage of these settings.

\section*{C}

\begin{itemize}

\item Visual consistency:  All menus, windows, and controls should have a
consistent look and feel to them.  Continues across all minor versions
of the interface.

\item Interface should respond quickly:  Any long running or inconsistent 
time operations should be run on a separate thread with an interface control
indicating the progress of the operation.

\item Operations should be stoppable:  Operations should be able to be
stopped and any changes should be rolled back to before the operation
was started.

\item Confirm dangerous operations:  Operations that cannot be undone 
should be confirmed with the user prior to being executed.

\item Do not crowd the screen:  Do not fill the screen with a large number
of controls or information.  This can confuse or overwhelm the user.

\item Do not punish users for mistakes:  Users will make mistakes.  Design
the interface to recover gracefully from these mistakes.  This can include
backups and saving or allowing users to edit information they have
invested time inputting into the system.

\item Reuse as much as possible  Reusing interface menus and components
makes the users more comfortable and increase the ease of use and memory 
retention of how to use the interface.

\item Use the users language:  The interface should be in the users
language.  This includes language related to the field or business it is
designed for.

\item Design for recognition and not recall:  The user should be able to
recognize how to use the interface and should not have to remember a set 
of complex steps to perform an operation.

\item Access to help and documentation:  The user should be able to access 
a searchable help system from within the interface.  The user should also
be able to access documentation outside of the interface.  

\item Help users understand and recover from errors:  Errors should be
express in a plain and understandable fashion.  It should indicate the 
problem and give possible solutions.

\end{itemize}

\section*{D}

The answer to this question is based on the paper small business storage 
management and gate control system.

The metaphor used for the storage site is based a 2D map of the facility.
The map shows the location of all storage units, parking spaces, and the 
gate.  Units on the map are color coded to give the manager a quick view
of the status of the facility.  Rented units and spaces would be in blue,
renters that have not paid are in red, unrentable units would be in orange,
and units that are ready to be rented are in green.  The color coding can
be configured.  Each unit is marked 
with a number and the size of the unit.  This screen also includes options
to change the color coding based on other attributes.  This can include 
unit sizes, how long a renter has had a unit, unit income over time, and
how often a unit is late on payments.

The site map can be useful for many other operations.  If a costumer wants 
to rent a unit the manager can quickly see on the map which units are ready
to be rented.  They can also view rentable units by size.  The manager can 
then click on a unit and select "Rent out unit" from the pop up menu.  From
my own experience many customers forget which unit number they have, but 
often can remember where their unit is on site and can find it on a map.  
This
can help the manager by keeping them from spending time searching for the
customer in the database.  The manager can select a unit on the site and 
from the menu and select several options.  These options include "Apply a
payment", "Move out", "Contact customer", "Change Rate", or "View details".



\end{document}

