\documentclass{letter}

\signature{Byron Heads}
\address{802 Lowell St. Apt 2\\ Ypsilanti, MI 48197 \\
517.376.0806 \\ bheads@emich.edu }

\begin{document}

\begin{letter}{ Stardock \\ 15090 Beck Rd. \\ Plymouth, MI 48170 }

\opening{Dear Hiring Manager}

I saw the posting for an Associate Game Developer on your web site.  I see this as a great opportunity to get an entry-level position into an industry I love.  As a Michigan native and an avid gamer, Stardock is on the top of my list of developers.  

Enclosed is a copy of my resume containing my education and experience, but you will find it lacks the ability to truly define my passion for creating games.  Video games are a medium with an infinite amount of expression and challenge all aspects of programming.  It is in this industry that I find my fulfilment as a programmer.

I am currently a Graduate Student working on maintaining computer labs and tutoring students.   I am in the last year as a masters student in computer science at Eastern Michigan University.  My thesis work is on real-time CPU bound ray-tracing.

I have experience I writing games in my free time.  I have experience in developing in C and C++ on Window, Linux and OS X.  I have worked with SDL, SFML, GLFW, OpenGL and DirectX on game projects in the past.  I have used GUI tool-kits such as Qt, wxWidgets, and .Net to create varying applications.

I look forward to hearing back from you in the future about an interview, thank you.

\closing{Sincerely,}
\encl{Cover Letter and Resume}
\end{letter}
\end{document}